Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-15 paper/main.tex
add TODOs
format figure; use newly generated image

\begin{itemize}
    \item This publication contains 5 things:
    \begin{enumerate}
        \item Methods of how to obtain data from microscopic images and perform parameter
            estimations; $\rightarrow$ also possibly across multiple generations
        \item Showcase the usefulness of the methods and the model by applying it to parameter
            estimation via microscopic images and multilayer formation
        \item Initial (incomplete) model of rod-shaped bacteria
        \item Generate 3D Images (masks)
        \item A python package which can do all of that and can be reused and extended
    \end{enumerate}
    \item Overall structure
    \begin{enumerate}
        \item users bring model + data
        \item this paper provides methods to do the parameter estimation + visualization
    \end{enumerate}
    \item For complete model, more work has to be done
    \begin{itemize}
        \item More aspects (see Review; put in discussion)
        \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
            important.
    \end{itemize}
\end{itemize}

@@

\begin{figure}
    \centering
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-004000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-006000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-008000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-calculations.png}%
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-distribution.png}
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}
    \caption{
        Progression of generated masks with estimated and exact positions.
        Comparison of masks across generations. Compare estimated with exact rod lengths.
    }
    \label{fig:benchmarking-extraction-algorithm}
\end{figure}

@@

@@

\begin{figure}
    \centering
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/fitting-methods/progressions-1.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/fitting-methods/progressions-2.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \linebreak
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/fitting-methods/progressions-3.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.3\textwidth]
        {docs/source/_static/fitting-methods/progressions-4.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}
    \caption{
        The figure shows two masks (A,B) which will be compared against each other.
        As we can see, image (B) contains more cells which has resulted from a division process that
        is not yet occurred in image (A).
        In (C) we can see the "naive" comparison, simply identifying non-matching colors.
        By assigning penalties to all highlighted pixels, we would not take into consideration the
        parental relationship.
        This approach is displayed in (D) where pixels of cells that are related have been colored
        in gray.
    }
    \label{fig:mask-difference-metric}
\end{figure}

>>>>

\begin{itemize}
    \item This publication contains 5 things:
    \begin{enumerate}
        \item Methods of how to obtain data from microscopic images and perform parameter
            estimations; $\rightarrow$ also possibly across multiple generations
        \item Showcase the usefulness of the methods and the model by applying it to parameter
            estimation via microscopic images and multilayer formation
        \item Initial (incomplete) model of rod-shaped bacteria
        \item Generate 3D Images (masks)
        \item A python package which can do all of that and can be reused and extended
    \end{enumerate}
    \item Overall structure
    \begin{enumerate}
        \item users bring model + data
        \item this paper provides methods to do the parameter estimation + visualization
    \end{enumerate}
    \item For complete model, more work has to be done
    \begin{itemize}
        \item More aspects (see Review; put in discussion)
        \item Note that we only care about \textit{E.Coli} and \textit{B.Subtilits}! This is
            important.
    \end{itemize}
    \begin{itemize}
        \item State of the Art
    \end{itemize}
\end{itemize}

@@

\begin{figure}
    \centering
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-004000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-006000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.32\textwidth]
        {docs/source/_static/fitting-methods/extract_positions-008000.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-calculations.png}%
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}%
    \begin{tikzonimage}[width=0.5\textwidth]
        {docs/source/_static/fitting-methods/displacement-distribution.png}
        \node at (0.03, 0.99)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
    \end{tikzonimage}
    \caption{
        Progression of generated masks with estimated and exact positions.
        Comparison of masks across generations. Compare estimated with exact rod lengths.
        "Iceberg plot"
    }
    \label{fig:benchmarking-extraction-algorithm}
\end{figure}

@@

\begin{itemize}
    \item Semantic Segmentation is more general
    \item Loss functions and some techniques can be reused
    \item Discuss which types of loss functions we could consider
    \item Possibly benchmark them later when doing the parameter estimation; compare loss functions
\end{itemize}

@@

\begin{figure}[H]
    \centering
    \begin{minipage}{0.5\textwidth}
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-1.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-2.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{B}};
    \end{tikzonimage}
    \linebreak
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-3.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{C}};
    \end{tikzonimage}%
    \hspace{0.01\textwidth}%
    \begin{tikzonimage}[width=0.48\textwidth]
        {docs/source/_static/fitting-methods/progressions-4.png}
        \node at (0.025, 0.975)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{D}};
    \end{tikzonimage}
    \end{minipage}%
    \begin{minipage}{0.49\textwidth}
        \begin{tikzonimage}[width=\textwidth]
            {docs/source/_static/fitting-methods/penalty-time-flow.png}%
            \node at (0, 0.975)[anchor=north west, draw, rectangle, minimum width=15pt, minimum height=15pt]{\textbf{E}};
        \end{tikzonimage}
    \end{minipage}
    \caption{
        The figure shows two masks (A,B) which will be compared against each other.
        As we can see, image (B) contains more cells which has resulted from a division process that
        is not yet occurred in image (A).
        In (C) we can see the "naive" comparison, simply identifying non-matching colors.
        By assigning penalties to all highlighted pixels, we would not take into consideration the
        parental relationship.
        This approach is displayed in (D) where pixels of cells that are related have been colored
        in gray.
    }
    \label{fig:mask-difference-metric}
\end{figure}
